---
title: "Survey variables analysis to produce Sanky Diagram"
author: "Ryan Kopper"
date: "5/22/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
## load libraries
library(dplyr)
#info sheet https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf
library(networkD3)
library(tidyverse)
library(readxl)
library(readr)

#Read in each sheet individually which creates a Tibble for each.

Relationships <- read_excel("/Users/ryankopper/Clark University/Morgan Ruelle - Ryan Kopper Research/Push Pull Systematic Review/Survey relationships 2020_05_19.xlsx", sheet = "Relationships")

Independent_recode <- read_excel("/Users/ryankopper/Clark University/Morgan Ruelle - Ryan Kopper Research/Push Pull Systematic Review/Survey relationships 2020_05_19.xlsx", sheet = "Independent recode")

Dependent_recode <- read_excel("/Users/ryankopper/Clark University/Morgan Ruelle - Ryan Kopper Research/Push Pull Systematic Review/Survey relationships 2020_05_19.xlsx", sheet = "Dependent recode")

```

Join independent/dependent variable of recode table to original,

https://www.infoworld.com/article/3454356/how-to-merge-data-in-r-using-r-merge-dplyr-or-datatable.html

The first step is some kind of join command 

the second one is using the joined columns to create a new matrix with counts.

alter long data to wide data with value as counts.

https://datacarpentry.org/R-ecology-lesson/03-dplyr.html
```{r}

joined_long <- left_join( Relationships, Independent_recode, by = "Independent_variable") %>% left_join(Dependent_recode, by = "Dependent_variable") %>%  select(Independent_variable_recode, Dependent_variable_recode) %>% group_by(Independent_variable_recode, Dependent_variable_recode) %>% summarise(count = n()) 

joined_matrix <- left_join( Relationships, Independent_recode, by = "Independent_variable") %>% left_join(Dependent_recode, by = "Dependent_variable") %>%  select(Independent_variable_recode, Dependent_variable_recode) %>% group_by(Independent_variable_recode, Dependent_variable_recode) %>% summarise(count = n()) %>% spread(key = Dependent_variable_recode, value = count) %>% replace(is.na(.), 0 ) 

# glimpse(joined_tibble)
# str(joined_tibble)

```

```{r}


```


```{r}
#Bubble plot

library(scales)
library(RColorBrewer)

ggplot(joined_long, aes(x = Dependent_variable_recode, y = Independent_variable_recode, color = count )) + geom_point(aes(size = count)) + theme(axis.text.x = element_text(angle = 90, vjust = .5 ,  hjust = 1)) + scale_x_discrete(labels = label_wrap(20)) + scale_y_discrete(labels = label_wrap(30)) 

#polish graph and note the oberservations

#change gradient to more distinct colors

#removed doubled count key
g1 <- ggplot(joined_long, aes(x = Dependent_variable_recode, y = Independent_variable_recode, color = count)) + geom_point(aes(size = count)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = .5 ,  hjust = 1)) + scale_x_discrete(labels = label_wrap(20)) + scale_y_discrete(labels = label_wrap(30)) + guides(size = FALSE) + xlab("Dependent variable") + ylab("Independent variable")

#COLOR OPTIONS

#change color to green

g1 +  scale_color_gradient(low = 'greenyellow', high = 'darkgreen') + guides(color = guide_legend(), size = guide_legend()) 

#change color to green
g1 + scale_color_gradient2(low = 'orange', mid = 'white', high = 'purple') + guides(size = FALSE) 

#look into fixing the x axis text labels to have multiple lines

ggplot(joined_long, aes(x = Dependent_variable_recode, y = Independent_variable_recode, color = count )) + geom_point(aes(size = count)) + theme(axis.text.x = element_text(angle = 90, vjust = .5 ,  hjust = 1)) + scale_x_discrete(labels = label_wrap(20)) + scale_y_discrete(labels = label_wrap(30)) 

```

EXTRA-------------------------------
Use DPLYR to bring together the "relationships" and "recording" sheets
 
```{r}
#Creates value for  each sheet
# sheets <- excel_sheets("/Users/ryankopper/Clark University/Morgan Ruelle - Ryan Kopper Research/Push Pull Systematic Review/Survey relationships 2020_05_19.xlsx")
# 
# #Maps sheets to the data
# #map_df() returns a data frame.
# f <- map_df(sheets, ~ read_excel("/Users/ryankopper/Clark University/Morgan Ruelle - Ryan Kopper Research/Push Pull Systematic Review/Survey relationships 2020_05_19.xlsx", sheet = .x))

#f = all the sheets combined.
```


SANKEY DIAGRAMS

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#SANKEY diagrams consist of three sets of elements: the nodes, the links, and the instructions which determine their positions

library(networkD3)

links2 <- read_excel("/Users/ryankopper/Clark University/Morgan Ruelle - Ryan Kopper Research/Push Pull Systematic Review/Survey relationships 2020_05_19.xlsx", sheet = "links")

nodes2 <- read_excel("/Users/ryankopper/Clark University/Morgan Ruelle - Ryan Kopper Research/Push Pull Systematic Review/Survey relationships 2020_05_19.xlsx", sheet = "nodes") 


```
https://www.youtube.com/watch?v=S6me1r6RI4I

Nodes
"networkD3" Package

https://towardsdatascience.com/using-networkd3-in-r-to-create-simple-and-clear-sankey-diagrams-48f8ba8a4ace


CREATING IT IN R
```{r}
#
# create nodes dataframe

IV <- unique(joined_long$Independent_variable_recode)
DV <- unique(joined_long$Dependent_variable_recode)
node <- data.frame(key = 0:30, variable = c(IV, DV))


#create links dataframe and rename 
links <- joined_long %>% inner_join(node,  by = c("Independent_variable_recode" = "variable")) %>% rename(IV_key = key) %>% 
  inner_join(node,  by = c("Dependent_variable_recode" = "variable")) %>% rename(DV_key = key)


sankeyNetwork(Links = links, Nodes = node, Source = 'IV_key', 
              Target = 'DV_key', Value = 'count', NodeID = "variable", 
              fontSize = 12, nodeWidth = 20, 
              fontFamily = "sans serif", iterations = 0)


```

```{r}
#LinkGroup
#when we bring in significant and non
#https://www.r-graph-gallery.com/322-custom-colours-in-sankey-diagram.html    


joined_long_sig <- Relationships %>% filter(Relationship == "significant") %>% left_join(Independent_recode, by = "Independent_variable") %>% left_join(Dependent_recode, by = "Dependent_variable") %>% select(Independent_variable_recode, Dependent_variable_recode) %>% group_by(Independent_variable_recode, Dependent_variable_recode) %>% summarise(significant = n()) %>% full_join(joined_long) %>% replace(is.na(.), 0 ) %>% mutate(significance = if_else(significant == 0, "no sig", if_else(significant == count, "sig", "mixed")))


# create nodes dataframe

IVsig <- unique(joined_long_sig$Independent_variable_recode)
DVsig <- unique(joined_long_sig$Dependent_variable_recode)
nodesig <- data.frame(key = 0:30, variable = c(IVsig, DVsig))


linkssig <- joined_long_sig %>% inner_join(nodesig,  by = c("Independent_variable_recode" = "variable")) %>% rename(IV_key = key) %>% 
  inner_join(nodesig,  by = c("Dependent_variable_recode" = "variable")) %>% rename(DV_key = key)


#create links dataframe and rename 

# Add a 'group' column to each node. Here I decide to put all of them in the same group to make them grey
node$group <- as.factor(c("group"))
 
# Give a color for each group:
# sig = green
# mixed yellow
# not sig = grey
my_color <- 'd3.scaleOrdinal() .domain(["sig","mixed","no sig"]) .range(["green", "orange", "grey"])'
 
# Make the Network

#no "no sig" grey lines are not appearing
sankeyNetwork(Links = linkssig,
              Nodes = nodesig,
              Source = 'IV_key', 
              Target = 'DV_key',
              Value = 'count',
              NodeID = "variable", 
              colourScale = my_color,
              LinkGroup = "significance",
              fontSize = 12, nodeWidth = 20, 
              fontFamily = "sans serif", iterations = 0)

#Attempt to change

sankeyNetwork(Links = linkssig,
              Nodes = nodesig,
              Source = "IV_key",
              Target = "DV_key",
              Value = "count",
              NodeID = "variable",
              LinkGroup = "significance")


??sankeyNetwork




```


